# STAT 305 Group work on Optimization

Please write the names and email addresses of your group members.

* Name / email (@wisc.edu only): Hadi Hashemi Nejad / mhashemineja@wisc.edu

# Part 1: One-dimensional optimization

Graph the object's altitude vs. time.

```{r}
rm(list=ls())
altitude = function(t, z0 = 100, g = 9.81, v0=55, m=80, c=15){
  return(z0 + (m/c)*(v0+m*g/c)*(1-exp(-c*t/m))-m*g*t/c)
}
curve(altitude, from = 0, to = 20, main = "Graph of altitude vs time")
```

Find the time at which the object strikes the ground.

```{r}
for(i in seq(11, 12, length.out=10000)){
  if(all.equal(altitude(t=i),0,tolerance=0.001)==TRUE){cat(i)}
}
```

Find the object's maximum height.

```{r}
optimize(altitude,lower=0,upper=20,maximum=TRUE)$objective
```

Find the time at which the object reaches its maximum height.

```{r}
optimize(altitude,lower=0,upper=20,maximum=TRUE)$maximum
```

# Part 2: Multi-dimensional optimization

Implement `gradient.descent()`.

```{r}
Z = function(x,y){
  return(7.9+0.13*x+0.21*y-0.05*x**2-0.016*y**2-0.007*x*y)
}
Z.prime = function(par){
  x=par[1]
  y=par[2]
  dZ.dx=0.13-0.1*x-0.007*y
  dZ.dy=0.21-0.032*y-0.007*x
  return(c(dZ.dx,dZ.dy))
}
gradient.descent = function(par, gr, gamma = .1, epsilon = .01, n = 100, verbose = FALSE, graph=FALSE, ...) {
  if(graph){
    x=seq(-10,10)
    y=seq(0,20)
    z=outer(x,y,Z)
    persp.out=persp(x,y,z,
    main="Distribution of pollutant concentration in a channel",
    zlab = "Concentration",
    theta = 45, phi = 45,
    col = "orange", shade = 0.4)
    }
  z.old = Z(par[1],par[2])
  for (i in seq_len(n)) {
    gradient = gr(par, ...)
    par = par + gamma * gradient
    z = Z(par[1], par[2])
    gradient.size = sum(abs(gradient))
    if (verbose) {
      cat(sep = "", "i = ", i, ", par = c(", paste(signif(par, 4), collapse = ","),
          "), gradient = c(", paste(signif(gradient, 4), collapse = ","),
          "), size = ", signif(gradient.size, 4), "\n")
    }
    if (graph) { # This block is not part of the algorithm. It just draws.
      points(trans3d(x = par[1], y = par[2], z = 0, pmat = persp.out), col = "blue", pch = 16)
      points(trans3d(x = par[1], y = par[2], z = z, pmat = persp.out), col = "red",  pch = 16)
      lines(trans3d(x = c(par[1], par[1]), y = c(par[2], par[2]), z = c(0, z), pmat = persp.out), col = "black")
    }
    if (gradient.size < epsilon || z.old > z) {
      break
    }
    z.old=z
  }
  if(graph){
    points(trans3d(x = par[1], y = par[2], z = z, pmat = persp.out), col = "yellow",  pch = 11)
  }
  return(par)
}

```
Graph the concentration.

Use `gradient.descent()` to find the peak.

Use `optim()` with `method=Nelder-Mead` to find the peak.

Use `optim()` with `method=BFGS` to find the peak.

How many calls did `optim()` make in each case? Which method would you
expect to be faster?

```{r}
#gradient descent concentration
g.d.c = gradient.descent(par = c(0, 0), gr = Z.prime, n = 1000, verbose = FALSE, graph = TRUE)
print(paste0("The peak concentration using gradient descent method is: ", Z(g.d.c[1],g.d.c[2])," where x=", g.d.c[1], " y=",g.d.c[2]))

Z.for.other.methods = function(par){
  x=par[1]
  y=par[2]
  return(7.9+0.13*x+0.21*y-0.05*x**2-0.016*y**2-0.007*x*y)
}

#Nelder-Mead concentration
n.m.c = optim(par=c(0,0), fn = Z.for.other.methods, method = "Nelder-Mead", control = list(fnscale = -1))
print(paste0("The peak concentration using Nelder-Mead method is: ", n.m.c$value," where x=", n.m.c$par[1], " y=",n.m.c$par[2]))

#BFGS concentration
bfgs.c = optim(par=c(0,0),gr=Z.prime,fn=Z.for.other.methods,method="BFGS",control = list(fnscale = -1))
print(paste0("The peak concentration using BFGS method is: ", bfgs.c$value," where x=", bfgs.c$par[1], " y=",bfgs.c$par[2]))

#counts
print(paste0("Nelder-Mead method made ",n.m.c$counts[[1]], " many calls to the concentration function"))
print(paste0("BFGS method made ",bfgs.c$counts[[1]], " many calls to the concentration function and ",bfgs.c$counts[[2]]," many calls to the concentration gradient"))

print("I expected BFGS to be the fastest")
```

